Automated Formal Synthesis of Provably Safe Digital Controllers for Continuous Plants

### 2019-07-09 automatica/response.tex

\vspace{1em}

{\bf Response: }

We have added the completeness threshold for each benchmark in the table. Note that the completeness threshold is dependent on the controller, 
and a controller with a smaller completeness threshold may exist.

The proof of Theorem 2 has been elaborated and made more comprehensive and detailed.

\todo[inline]{Elizabeth: add completeness thresholds to table}

@@

\vspace{2em}

{\bf Comment: }
{\itshape I would like to see some additional discussion and/or comparison to hybrid systems reachability. While the focus of this work is on lower-level
software synthesis, if the approach is to be believed in the context of closed-loop systems satisfying specifications, then I would like some
discussion/comparison to that vast existing literature. Specifically, I presume the reason the benchmarks are not scalable beyond ~n=4 dimensionality
(as the n=5 benchmark appears to timeout), which is trivially small at this stage for linear systems (e.g., given a recent HSCC'19 paper on verifying
$R^{billion}$ linear systems, or at least n=100 to n=1000 with standard approaches like SpaceEx), is due to overapproximation error growth with the
interval representation of plant behaviors, whereas most scalable approaches utilize better data structures for representing geometric sets
(zonotopes, support functions, star sets, Taylor models, etc.). This likewise limits the possible applicability of the
approach for nonlinear plants, although there are a host of technical issues in that potential extension. }

\vspace{1em}
{\bf Response: }

We have reviewed relevant literature on controller synthesis based on finite abstractions, cf. Belta and Tabuada.

Literature on reachability analysis tackles a different problem, namely one of verification, rather than one of control synthesis (which this work focuses on).

Admittedly reachability analysis tool scale quite well (as the reviewer points out), however this is under clear assumptions on model dynamics.

One might argue that reachability analysis tools could be embedded within the CEGIS framework, however they currently lack full soundness: namely they are algorithmically sound, but not numerically sound. On the contrary our work is sound on both sides. (Shall we mention that indeed we do reachability analysis via Axelerator in CAV17?)

@@

\vspace{2em}
{\bf Comment: }
{\itshape Finally, while checking the $2^n$ vertices
are sufficient (theorem 1), I doubt this is necessary, as e.g. the star set approaches avoid this blow-up that would also limit the dimensionality
the approach can work toward.}
\vspace{1em}

\rev{It is probably true that the check on all vertices might not be necessary in general, that that this check might be simplified in special cases or under special conditions. The check however, as it currently is implemented, does not represent the bottleneck of our overall method, so we can safely rely on its current form, as it stands. }

@@

The same theorem is also present in [1]. The proof, upon careful inspection, is not a proof of the statement given in the theorem. It only proves
that the reachable set is the convex set obtained by the trajectories of the corner states. However, it does not prove that if the trajectories are
safe, their convex combination is also safe. In fact, as given in the above counterexample, the vertices of polytope are safe does not mean that the
convex combination is safe.

I think this bug needs to be fixed in the BMC part of the verifier. Instead of checking the individual trajectories, you should in fact compute the
convex combination of trajectories (reachable set) and then perform the verification.}

\vspace{1em}
{\bf Response: }
\todo[inline]{Alessandro: Response to this (Theorem 1)}
\vspace{2em}

\rev{
Many thanks for this note. 
Notice that in the submitted manuscript we had focused on special shapes of initial and safe sets, namely (convex) hyper-boxes, which included the origin. 
In the amended draft, we have better clarified these assumptions, and 
we have emphasised that the initial set is assumed to lie within the safe set: indeed, the contrary would trivially result in an unsafe condition. 
In conclusion, recall that the sets are hyper-boxes, thus convex, and that in view of the linear dynamics of the considered models, in practice both are assumed to contain the origin.  }

@@

\section{Response to Reviewer (4)}

{\bf Comment:  } {\itshape The safety property in Eqn 3 focuses on interval constraints. One can generate auxiliary variables such that range constraints over them can encode
more complex safety properties. It would be nice to bring this up in the discussion after equation 3 that $x_i$ need not be just $i$-th component of the
state.  Can these also incorporate temporal properties (say STL properties with bounded time-horizon) without having to unroll the transition and
create auxiliary variables that track history explicitly?}

\todo[inline]{Eli: can we generalise according to the suggestion of the reviewer?}

@@

\vspace{1em}
{\bf Response: }
We have moved these figures to [] 
\todo[inline]{move figure}
\vspace{2em}

{\bf Comment: } {\itshape
Some other relevant applications of CEGIS to CPS/dynamical systems include: Synthesizing switching logic for
safety and dwell-time requirements in ICCPS, 2010; Synthesis of optimal switching logic for hybrid systems in EMSOFT, 2011, Synthesis of Optimal
Fixed-Point Implementation of Numerical Software Routines in NSV'13, Bridging boolean and quantitative synthesis using smoothed proof search,
POPL'14, Synthesis of fixed-point programs in EMSOFT'13}

\vspace{1em}
{\bf Response: }
We have added a discussion about these applications and results to the related work as follows:
\todo[inline]{add discussion}

\vspace{2em}
{\bf Comment: } {\itshape Could you add one example to the end of Section 4 to show $R_1$, $R_2$, $R_3$, $R_4$ for a simple LTI system?

\todo[inline]{Eli: can you clarify?}

\vspace{2em}
{\bf Comment: } {\itshape 
Section 4 is one the most interesting parts
of the paper and its current presentation is a bit sketchy. What are $m_{ij}$?} }

\todo[inline]{Eli: can we elaborate?}

### 2019-07-10 automatica/response.tex

\vspace{1em}

{\bf Response: }

This journal submission encompasses and summarises the mentioned conference articles (CAV, HSCC, ASE) or workshop abstract (SYNTH). 
It is a more mature, complete, and comprehensive presentation of the conference material. 
Of these conference articles, the CAV/HSCC contributions presented the theory underlying the approach, 
whereas the ASE is a tool paper. 
The application of the CEGIS architecture to the synthesis of digital controllers for continuous plants is indeed a novel contribution of this line of
work.

\vspace{2em}
{\bf Comment: } {\itshape The results are evaluated on about a dozen and a half examples from the controls literature all with relatively small
dimensionality (2 to 5 real variables in the continuous state space), and while the authors claim the software artifacts and benchmarks are available
for evaluation, when this reviewer browsed to the link (http://www.cprover.org/DSSynth/), a page with "TBD" was shown. I did a little digging and
eventually found what I presume is the same / similar artifacts here: https://github.com/ssvlab/dsverifier and/or http://dsverifier.org/}

@@

\rev{
Many thanks for this note. 
Notice that in the submitted manuscript we had focused on special shapes of initial and safe sets, namely (convex) hyper-boxes, which included the
origin. 
In the amended draft, we have better clarified these assumptions, and 
we have emphasised that the initial set is assumed to lie within the safe set: indeed, the contrary would trivially result in an unsafe condition. 
In conclusion, recall that the sets are hyper-boxes, thus convex, and that in view of the linear dynamics of the considered models, in practice both
are assumed to contain the origin.  }

\section{Response to Reviewer (4)}

{\bf Comment:  } {\itshape The safety property in Eqn 3 focuses on interval constraints. One can generate auxiliary variables such that range
constraints over them can encode
more complex safety properties. It would be nice to bring this up in the discussion after equation 3 that $x_i$ need not be just $i$-th component of
the
state.  Can these also incorporate temporal properties (say STL properties with bounded time-horizon) without having to unroll the transition and
create auxiliary variables that track history explicitly?}

\vspace{1em}
{\bf Response:}
 It is true that we would be able to encode more complex properties by introducing auxiliary variables; however theorem 2 is dependent on the safe set being convex, and 
so this proof would not necessarily hold with more complex properties. It would be possible to construct time-bounded properties as automata that track history explicitly, but I can't see how
it would be possible to avoid unwinding the transition relation using this method. In the case of the example given by the reviewer, an STL property with bounded time-horizon, unwinding
the transition relation would not pose much of a problem anyway as a time-bounded property would only require unwinding up to the time-bound.

%\todo[inline]{Eli: can we generalise according to the suggestion of the reviewer?}

@@

\vspace{2em}
{\bf Comment: } {\itshape The paper would benefit with a restructuring of the discussion. Figure 2 and 3 will be more useful
earlier in the paper (Section 1 or 3).}

\vspace{1em}
{\bf Response: }
We have moved these figures to section 1 and incorporated them into the introduction. We are still referring back to these figures from section 5.

### 2019-07-10 automatica/paper.tex
move figures

%% offer three modes of
%% operation that address different aspects of the problem.
%% {\em The first  approach} uses a transfer-function based analysis that
%% solves the problem of robust stability for closed loop systems for a
%% collection of unstable plants.  Inthis case, there is no safety
%% specification since we are working in the frequency domain.
\ifx\axelerator
\emph{The first approach} 
\else
Our approach
\fi
uses a multi-staged technique that starts by devising a
digital controller that stabilizes the plant model while remaining safe for a
pre-selected time horizon and a single initial state; then, it verifies
unbounded-time safety by unfolding the model dynamics, considering the 
full set of initial states, and checking a \emph{completeness
threshold}~\cite{DBLP:conf/vmcai/KroeningS03}, i.e., 
the number of stages required to sufficiently unwind the closed-loop model, 
such that the safety boundaries \rev{(assumed to be a compact set)} are not violated for any larger number of iterations.

@@

\begin{figure*}[htb]
\scriptsize\centering
\begin{tikzpicture}[scale=0.3,->,>=stealth',shorten >=.2pt,auto, semithick, initial text=, ampersand replacement=\&,]
  \matrix[nodes={draw, fill=none, shape=rectangle, minimum height=.2cm, minimum width=.2cm, align=center},
          row sep=.6cm, column sep=.9cm] {
   \coordinate (aux1);
   \& \coordinate (aux2);
   \&;\\
   \coordinate (aux3);
   \& \coordinate (aux4);
   \&;\\
   \coordinate (aux5);
   \& \coordinate (aux6);
   \&;\\
   \node[minimum width=1.5cm, minimum height=0.6cm, fill=gray!20] (synth) {{\sc 1.Synthesize}};
   \&
   complexnode/.pic={ 
     \node[rectangle,draw,dotted,
	minimum width=6cm,
	minimum height=1cm,
        pattern=north west lines, pattern color=gray!20,
	label={\sc ~~~~~~~~~~~~Verify},] (verif) {};
     \node[minimum width=1cm, minimum height=0.6cm, fill=gray!20] (verif1) at ([xshift=-2cm]verif.center) {{\sc 2.Safety}};
     \node[minimum width=1cm, minimum height=0.6cm, fill=gray!20] (verif2) at ([xshift=0cm]verif.center) {{\sc 3.Precision}};
     \node[minimum width=1cm, minimum height=0.6cm, fill=gray!20] (verif3) at ([xshift=2cm]verif.center) {{\sc 4.Complete}};
     %\node[minimum width=1cm, minimum height=0.6cm, fill=gray!20] (verif4) at ([xshift=3.1cm]verif.center) {{\sc 5.Sampling}};
   } 
   \& \node[ellipse, fill=gray!20] (done) {{\sc Done}};\\
   \& \\
   \node[minimum height=0cm] (gp) {\sf Program Search};
   \&
   complexnode/.pic={ 
     \coordinate (aux);
   \node (bmc) at ([xshift=-2cm]aux.center) {\sf BMC-based \\ \sf Verifier};
   \node (fp)  at ([xshift=0cm]aux.center) {\sf Fixed-point \\ \sf Arithmetic\\ \sf Verifier};
   \node (sv)  at ([xshift=2cm]aux.center) {\sf Completeness\\ \sf Verifier};
   %\node (cv)  at ([xshift=3.1cm]aux.center) {\sf Sampling\\Verifier};
   }   
    \\
  };

   \path
    ([yshift=2em]synth.east) edge node[xshift=-0.5em,align=center] {$K$} ([yshift=2em]verif1.west)
    ([yshift=-2em]verif1.west) edge node {C-ex} ([yshift=-2em]synth.east)
    ([xshift=-5em]fp.north) edge node[align=center]  {T/F} ([xshift=-5em]verif2.south)
    ([xshift=-5em]sv.north) edge node[align=center]  {T/F} ([xshift=-5em]verif3.south)
    %([xshift=-5em]cv.north) edge node[align=center]  {T/F} ([xshift=-5em]verif4.south)
    ([xshift=5em]verif1.south) edge node[align=center] {$K$} ([xshift=5em]bmc.north)
    ([xshift=5em]verif2.south) edge node[align=center] {$K$} ([xshift=5em]fp.north)
    ([xshift=5em]verif3.south) edge node[align=center] {$K$} ([xshift=5em]sv.north)
    %([xshift=5em]verif4.south) edge node[align=center] {$K$} ([xshift=5em]cv.north)
    ([xshift=-5em]bmc.north) edge node[align=center]  {UNSAT/\\model} ([xshift=-5em]verif1.south)
    (verif) edge node {PASS} (done)
    ([xshift=5em]synth.south) edge node[align=center] {Inputs} ([xshift=5em]gp.north)
    ([xshift=-5em]gp.north) edge node[align=center] {UNSAT/\\$K$} ([xshift=-5em]synth.south)
    (aux3) edge (synth.north);
   \path[-]
   (verif2.north) edge node[align=center] {} ([xshift=0cm]aux6)
   ([xshift=0cm]aux6) edge node[align=center] {Increase Precision} (aux5)
   (verif3.north) edge node[align=center] {} ([xshift=6.7cm]aux4)
   ([xshift=6.7cm]aux4) edge node[align=center] {Increase Unfolding Bound} (aux3);
   %(verif4.north) edge node[align=center] {} ([xshift=10.5cm]aux2)
   %([xshift=10.5cm]aux2) edge node[align=center] {Increase Sampling Rate} (aux1);
\end{tikzpicture}
\caption{CEGIS with multi-staged verification for digital controller synthesis}
\label{fig:CEGIS-precision-increment}
\end{figure*}

@@

>>>>

%% offer three modes of
%% operation that address different aspects of the problem.
%% {\em The first  approach} uses a transfer-function based analysis that
%% solves the problem of robust stability for closed loop systems for a
%% collection of unstable plants.  Inthis case, there is no safety
%% specification since we are working in the frequency domain.
\ifx\axelerator
\emph{The first approach} 
\else
\begin{figure*}[htb]
\scriptsize\centering
\begin{tikzpicture}[scale=0.3,->,>=stealth',shorten >=.2pt,auto, semithick, initial text=, ampersand replacement=\&,]
  \matrix[nodes={draw, fill=none, shape=rectangle, minimum height=.2cm, minimum width=.2cm, align=center},
          row sep=.6cm, column sep=.9cm] {
   \coordinate (aux1);
   \& \coordinate (aux2);
   \&;\\
   \coordinate (aux3);
   \& \coordinate (aux4);
   \&;\\
   \coordinate (aux5);
   \& \coordinate (aux6);
   \&;\\
   \node[minimum width=1.5cm, minimum height=0.6cm, fill=gray!20] (synth) {{\sc 1.Synthesize}};
   \&
   complexnode/.pic={ 
     \node[rectangle,draw,dotted,
	minimum width=6cm,
	minimum height=1cm,
        pattern=north west lines, pattern color=gray!20,
	label={\sc ~~~~~~~~~~~~Verify},] (verif) {};
     \node[minimum width=1cm, minimum height=0.6cm, fill=gray!20] (verif1) at ([xshift=-2cm]verif.center) {{\sc 2.Safety}};
     \node[minimum width=1cm, minimum height=0.6cm, fill=gray!20] (verif2) at ([xshift=0cm]verif.center) {{\sc 3.Precision}};
     \node[minimum width=1cm, minimum height=0.6cm, fill=gray!20] (verif3) at ([xshift=2cm]verif.center) {{\sc 4.Complete}};
     %\node[minimum width=1cm, minimum height=0.6cm, fill=gray!20] (verif4) at ([xshift=3.1cm]verif.center) {{\sc 5.Sampling}};
   } 
   \& \node[ellipse, fill=gray!20] (done) {{\sc Done}};\\
   \& \\
   \node[minimum height=0cm] (gp) {\sf Program Search};
   \&
   complexnode/.pic={ 
     \coordinate (aux);
   \node (bmc) at ([xshift=-2cm]aux.center) {\sf BMC-based \\ \sf Verifier};
   \node (fp)  at ([xshift=0cm]aux.center) {\sf Fixed-point \\ \sf Arithmetic\\ \sf Verifier};
   \node (sv)  at ([xshift=2cm]aux.center) {\sf Completeness\\ \sf Verifier};
   %\node (cv)  at ([xshift=3.1cm]aux.center) {\sf Sampling\\Verifier};
   }   
    \\
  };

   \path
    ([yshift=2em]synth.east) edge node[xshift=-0.5em,align=center] {$K$} ([yshift=2em]verif1.west)
    ([yshift=-2em]verif1.west) edge node {C-ex} ([yshift=-2em]synth.east)
    ([xshift=-5em]fp.north) edge node[align=center]  {T/F} ([xshift=-5em]verif2.south)
    ([xshift=-5em]sv.north) edge node[align=center]  {T/F} ([xshift=-5em]verif3.south)
    %([xshift=-5em]cv.north) edge node[align=center]  {T/F} ([xshift=-5em]verif4.south)
    ([xshift=5em]verif1.south) edge node[align=center] {$K$} ([xshift=5em]bmc.north)
    ([xshift=5em]verif2.south) edge node[align=center] {$K$} ([xshift=5em]fp.north)
    ([xshift=5em]verif3.south) edge node[align=center] {$K$} ([xshift=5em]sv.north)
    %([xshift=5em]verif4.south) edge node[align=center] {$K$} ([xshift=5em]cv.north)
    ([xshift=-5em]bmc.north) edge node[align=center]  {UNSAT/\\model} ([xshift=-5em]verif1.south)
    (verif) edge node {PASS} (done)
    ([xshift=5em]synth.south) edge node[align=center] {Inputs} ([xshift=5em]gp.north)
    ([xshift=-5em]gp.north) edge node[align=center] {UNSAT/\\$K$} ([xshift=-5em]synth.south)
    (aux3) edge (synth.north);
   \path[-]
   (verif2.north) edge node[align=center] {} ([xshift=0cm]aux6)
   ([xshift=0cm]aux6) edge node[align=center] {Increase Precision} (aux5)
   (verif3.north) edge node[align=center] {} ([xshift=6.7cm]aux4)
   ([xshift=6.7cm]aux4) edge node[align=center] {Increase Unfolding Bound} (aux3);
   %(verif4.north) edge node[align=center] {} ([xshift=10.5cm]aux2)
   %([xshift=10.5cm]aux2) edge node[align=center] {Increase Sampling Rate} (aux1);
\end{tikzpicture}
\caption{CEGIS with multi-staged verification for digital controller synthesis}
\label{fig:CEGIS-precision-increment}
\end{figure*}
Our approach
\fi
uses a multi-staged technique\rev{, shown in Figure~\ref{fig:CEGIS-precision-increment}}, that starts by devising a
digital controller that stabilizes the plant model while remaining safe for a
pre-selected time horizon and a single initial state; then, it verifies
unbounded-time safety by unfolding the model dynamics, considering the 
full set of initial states, and checking a \emph{completeness
threshold}~\cite{DBLP:conf/vmcai/KroeningS03}, i.e., 
the number of stages required to sufficiently unwind the closed-loop model 
such that the safety boundaries \rev{(assumed to be a compact set)} are not violated for any 
larger number of iterations\rev{, illustrated in Figure~\ref{fig:ct}}.

@@

%\begin{figure*}[htb]
%\scriptsize\centering
%\begin{tikzpicture}[scale=0.3,->,>=stealth',shorten >=.2pt,auto, semithick, initial text=, ampersand replacement=\&,]
%  \matrix[nodes={draw, fill=none, shape=rectangle, minimum height=.2cm, minimum width=.2cm, align=center},
%          row sep=.6cm, column sep=.9cm] {
%   \coordinate (aux1);
%   \& \coordinate (aux2);
%   \&;\\
%   \coordinate (aux3);
%   \& \coordinate (aux4);
%   \&;\\
%   \coordinate (aux5);
%   \& \coordinate (aux6);
%   \&;\\
%   \node[minimum width=1.5cm, minimum height=0.6cm, fill=gray!20] (synth) {{\sc 1.Synthesize}};
%   \&
%   complexnode/.pic={ 
%     \node[rectangle,draw,dotted,
%	minimum width=6cm,
%	minimum height=1cm,
%        pattern=north west lines, pattern color=gray!20,
%	label={\sc ~~~~~~~~~~~~Verify},] (verif) {};
%     \node[minimum width=1cm, minimum height=0.6cm, fill=gray!20] (verif1) at ([xshift=-2cm]verif.center) {{\sc 2.Safety}};
%     \node[minimum width=1cm, minimum height=0.6cm, fill=gray!20] (verif2) at ([xshift=0cm]verif.center) {{\sc 3.Precision}};
%     \node[minimum width=1cm, minimum height=0.6cm, fill=gray!20] (verif3) at ([xshift=2cm]verif.center) {{\sc 4.Complete}};
%     %\node[minimum width=1cm, minimum height=0.6cm, fill=gray!20] (verif4) at ([xshift=3.1cm]verif.center) {{\sc 5.Sampling}};
%   } 
%   \& \node[ellipse, fill=gray!20] (done) {{\sc Done}};\\
%   \& \\
%   \node[minimum height=0cm] (gp) {\sf Program Search};
%   \&
%   complexnode/.pic={ 
%     \coordinate (aux);
%   \node (bmc) at ([xshift=-2cm]aux.center) {\sf BMC-based \\ \sf Verifier};
%   \node (fp)  at ([xshift=0cm]aux.center) {\sf Fixed-point \\ \sf Arithmetic\\ \sf Verifier};
%   \node (sv)  at ([xshift=2cm]aux.center) {\sf Completeness\\ \sf Verifier};
%   %\node (cv)  at ([xshift=3.1cm]aux.center) {\sf Sampling\\Verifier};
%   }   
%    \\
%  };
%
%   \path
%    ([yshift=2em]synth.east) edge node[xshift=-0.5em,align=center] {$K$} ([yshift=2em]verif1.west)
%    ([yshift=-2em]verif1.west) edge node {C-ex} ([yshift=-2em]synth.east)
%    ([xshift=-5em]fp.north) edge node[align=center]  {T/F} ([xshift=-5em]verif2.south)
%    ([xshift=-5em]sv.north) edge node[align=center]  {T/F} ([xshift=-5em]verif3.south)
%    %([xshift=-5em]cv.north) edge node[align=center]  {T/F} ([xshift=-5em]verif4.south)
%    ([xshift=5em]verif1.south) edge node[align=center] {$K$} ([xshift=5em]bmc.north)
%    ([xshift=5em]verif2.south) edge node[align=center] {$K$} ([xshift=5em]fp.north)
%    ([xshift=5em]verif3.south) edge node[align=center] {$K$} ([xshift=5em]sv.north)
%    %([xshift=5em]verif4.south) edge node[align=center] {$K$} ([xshift=5em]cv.north)
%    ([xshift=-5em]bmc.north) edge node[align=center]  {UNSAT/\\model} ([xshift=-5em]verif1.south)
%    (verif) edge node {PASS} (done)
%    ([xshift=5em]synth.south) edge node[align=center] {Inputs} ([xshift=5em]gp.north)
%    ([xshift=-5em]gp.north) edge node[align=center] {UNSAT/\\$K$} ([xshift=-5em]synth.south)
%    (aux3) edge (synth.north);
%   \path[-]
%   (verif2.north) edge node[align=center] {} ([xshift=0cm]aux6)
%   ([xshift=0cm]aux6) edge node[align=center] {Increase Precision} (aux5)
%   (verif3.north) edge node[align=center] {} ([xshift=6.7cm]aux4)
%   ([xshift=6.7cm]aux4) edge node[align=center] {Increase Unfolding Bound} (aux3);
%   %(verif4.north) edge node[align=center] {} ([xshift=10.5cm]aux2)
%   %([xshift=10.5cm]aux2) edge node[align=center] {Increase Sampling Rate} (aux1);
%\end{tikzpicture}
%\caption{CEGIS with multi-staged verification for digital controller synthesis}
%\label{fig:CEGIS-precision-increment}
%\end{figure*}

@@

%\begin{figure}[t]
%\centering
%\includegraphics[width=.7\columnwidth]{figures/ct.eps}
%%\vspace{0.1cm}
%\caption{Completeness threshold for multi-staged verification (quantity $T_s$ is the time discretization step) }
%\label{fig:ct}
%\end{figure}
